% ash with student-t likelihood
% Sen Tian
% 05/26/2014

```{r, include=FALSE}
library(ashr)
```

```{r setup, include=FALSE}
# set global chunk options
opts_chunk$set(warning=FALSE)
dep_auto()
```
 
# Examples of using the student-t likelihood
- If 'df' option is specified in function 'ash', the student-t likelihood is applied.
- student-t likelihood with normal mixture is not yet implemented

### Example1.1: density estimations with two different likelihoods
```{r, echo=TRUE}
  simdata = function(n, nnull, altmean, altsd, betasd, df) {
    null = c(rep(1, nnull), rep(0, n - nnull))
    beta = c(rep(0, nnull), rnorm(n - nnull, altmean, altsd))
    betahat =  beta + betasd * rt(n,df)
    return(list(null = null, beta = beta, betahat = betahat, betasd = betasd))
  }
  
  ss = simdata(1000, 800, 0, 2, 1, 1.5)
  beta.ash.u_n = ash(ss$betahat, ss$betasd, mixcompdist="uniform",method="shrink", cxx=FALSE)
  beta.ash.u_t1 = ash(ss$betahat, ss$betasd, mixcompdist="uniform",method="shrink", cxx=FALSE,df=1.5)
  beta.ash.u_t2 = ash(ss$betahat, ss$betasd, mixcompdist="uniform",method="shrink", cxx=FALSE,df=5)
  beta.ash.u_t1n = ash(ss$betahat, ss$betasd, mixcompdist="uniform",method="fdr", cxx=FALSE,df=1.5)
  
  # plot the cdf
  x = seq(-6, 6, length = 1000)
  plot(ecdf(ss$beta), main="CDF, Uniform mixture",xlim=c(-6,6))
  lines(x,cdf.ash(beta.ash.u_n,x)$y,col=2,lwd=2)
  lines(x,cdf.ash(beta.ash.u_t1,x)$y,col=3,lwd=2)
  lines(x,cdf.ash(beta.ash.u_t2,x)$y,col=4,lwd=2)
  lines(x,cdf.ash(beta.ash.u_t1n,x)$y,col=5,lwd=2)
  legend("topleft",legend = c("normal-lik","t-lik,df=1.5","t-lik, df=5","t-lik, df=1.5, fdr"), 
         col = c(2,3,4,5), lwd = c(2,2,2,2))
```

### Example1.2: density estimation with student-t likelihood and normal likelihood with quasi df
```{r, echo=TRUE}
  effective.effect=function(betahat,se,df){
    p = pt(betahat/se,df)
    qnorm(p,sd=se)
  }
  ss$betahat = effective.effect(ss$betahat,ss$betasd,1.5)
  beta.ash.u_n1 =  ash(ss$betahat, ss$betasd, mixcompdist="uniform",method="shrink", cxx=FALSE)
  plot(ecdf(ss$beta), main="CDF, Uniform mixture",xlim=c(-6,6))
  lines(x,cdf.ash(beta.ash.u_n1,x)$y,col=2,lwd=2)
  lines(x,cdf.ash(beta.ash.u_t1,x)$y,col=3,lwd=2)
  legend("topleft",legend = c("normal-lik with quasi df=1.5","t-lik,df=1.5"), 
         col = c(2,3), lwd = c(2,2))
```

