```{r}
  setwd("~/Documents/git/ash/paper/Rcode")
  require(ashr)
  require(qvalue)

  altcol="cyan" #colors to use
  nullcol="blue" 
  nc=40 #number of bins in histograms
  ncz = 100 # number of bins in z score histograms
  
  hh = read.table("../../data/nejm_brca_release_edit.csv",sep=",",skip=3)
  subset = apply(hh, 1, max)<20
  hh = hh[subset,]

  labs = read.table("../../data/nejm_brca_release_edit.csv",sep=",",skip=1,nrows=1)
  labs = 1*(labs=="BRCA1") + 2 * (labs=="BRCA2") 

  hh.betahat = apply(hh[,labs==1],1,mean) - apply(hh[,labs==2],1,mean)
  n1 = sum(labs==1)
  n2 = sum(labs==2)
  hh.sebetahat = sqrt(apply(hh[,labs==1],1,var)/n1 + apply(hh[,labs==2],1,var)/n2)
  hh.zscore = hh.betahat/hh.sebetahat
  hh.pval = pchisq(hh.zscore^2,df=1,lower.tail=F)
  pdf("figures/FDR_illus.pdf")
  nc=40 #number of bins in histogram
  hh.hist=hist(hh.pval,freq=FALSE,xlab="p value",main="Distribution of p values for Hedenfalk et al data",nclass=nc,col=altcol)

  hh.q = qvalue(hh.pval)
  abline(h=hh.q$pi0,col=nullcol,lwd=2)
  
  hh.hist$density=rep(hh.q$pi0,length(hh.hist$density))  
  #hh.hist$counts=rep(hh.q$pi0*length(hh.pval)/nc,length(hh.hist$counts)) 
  plot(hh.hist,add=TRUE,col=nullcol,freq=FALSE)
  
  abline(v=0.1,lwd=2,col=2)
  
  text(0.05,1.1,labels="A",col=2,cex=1.7)  
  text(0.05,0.4,labels="B",col=2,cex=1.7)  
  text(0.4,3,labels=paste0("FDR = B/(A+B) =  ",round(hh.q$pi0*0.1*length(hh.pval)/sum(hh.pval<0.1),2)),cex=1.7)
  dev.off()
```

Now I wanted to investigate what distribution for z values under the alternative the standard approach implicitly corresponds to.
We use three different approaches.
The first, fdrtool, we use to directly model the p values. fdrtool assumes
that the density of p values under the alternative is monotonically decreasing away from 0, and fits a non-parametric maximum likelihood distribution under this asssumption.

The other two approaches, locfdr and mixfdr, we use to directly model the z scores.

In each case, the distribution of z scores under the alternative is implied by a) the overall distribution of z scores, f(z); and b) the localfdr for each z score. Specifically $f_1(z)=(1-lfdr(z)) * f(z)$.

Note that mixfdr does not make the zero assumption, and actually models
the z score errors, so its estimated distribution for f1 is at least theoretically possible. However, we still view thie bimodal distribution as generally implausible.

Our conclusion is that making the zero assumption leads to distributions of the z scores under the
alternative that are generally implausible.
```{r}
  source("nullalthist.R")
  
  pdf("figures/zhist.pdf")
  par(mfcol=c(3,1))
  require(fdrtool)
  hh.fdrtool = fdrtool(hh.pval,statistic="pvalue",plot=FALSE)
  nullalthist(hh.zscore,hh.fdrtool$lfdr,main="fdrtool")
  
  require(locfdr)
  hh.locfdr = locfdr(hh.zscore,nulltype=0,plot=0)
  nullalthist(hh.zscore,hh.locfdr$fdr,main="locfdr")
 
  require(mixfdr)
  hh.mixfdr = mixFdr(hh.zscore,noiseSD=1,theonull=TRUE,plot=FALSE)
  nullalthist(hh.zscore,hh.mixfdr$fdr,main="mixfdr")
  dev.off()
  
  par(mfcol=c(1,1))
  pdf("figures/zhist_fdrtool.pdf")
  require(fdrtool)
  hh.fdrtool = fdrtool(hh.pval,statistic="pvalue",plot=FALSE)
  nullalthist(hh.zscore,hh.fdrtool$lfdr,main="fdrtool")
  dev.off()
  
  pdf("figures/zhist_locfdr.pdf")
  require(locfdr)
  hh.locfdr = locfdr(hh.zscore,nulltype=0,plot=0)
  nullalthist(hh.zscore,hh.locfdr$fdr,main="locfdr")
  dev.off()
  
  pdf("figures/zhist_mixfdr.pdf")
  require(mixfdr)
  hh.mixfdr = mixFdr(hh.zscore,noiseSD=1,theonull=TRUE,plot=FALSE)
  nullalthist(hh.zscore,hh.mixfdr$fdr,main="mixfdr")
  dev.off()
  
  
```

```{r}
  hh.ashz = ash(hh.zscore,1,usePointMass = TRUE, VB=TRUE, prior="nullbiased")

  #now try a different prior to check robustness
  mixsd = c(0,0.01,0.02,0.04,0.08,0.16,0.32,0.64,1.28,2.56,5.12,10.24)
  prior = rep(1,length(mixsd))
  prior[1] = 100
  hh.ashz2 = ash(hh.zscore,1,mixsd=mixsd,prior=prior,VB=TRUE)

  pdf("figures/zhist_ash.pdf")
  nullalthist(hh.zscore,hh.ashz$ZeroProb,main="ash")
  dev.off()

```

Compare pi0 estimates
```{r}
  c(hh.fdrtool$param[3],
  hh.locfdr$fp0[1,3],
  hh.mixfdr$pi[1],
  hh.ashz$fitted.g$pi[1])
```

Number signficant at 0.05 localfdr
```{r}
  c(sum(hh.fdrtool$lfdr<0.05),
  sum(hh.locfdr$fdr<0.05),
  sum(hh.mixfdr$fdr<0.05),
  sum(hh.ashz$ZeroProb<0.05))
```





